Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex
Updated

@@

@@

	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} \nonumber\\
	&\qquad-\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\

@@

	-\frac{f_t}{b}\VARIS,\nonumber\\

@@

	-\frac{f_t}{b}\VARIS, \label{eq:20}

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\{\Psi_t\}, \\
& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi} \\
& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi}.
\end{align*}
The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.

Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ we get:
\begin{align}
\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
 \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 %
 &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
 &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 %
 &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
 &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 %
 &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
 &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
 %
 &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
 \end{align}
where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
Next, summing over epochs:
\begin{align}
\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
%
&\leq
\frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
 \label{eq:23}\\
%
&\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
\end{align}
where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.

Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
\begin{align}
\EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
&\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
\end{align}
\end{proof}

>>>>

\begin{equation}\label{E:lyapunov.function}

@@

\end{equation}

@@

	&\qquad-\frac{\alpha_t}{2N}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right)\Vars{g(\cdot\vert\tilde{\vtheta}^s)} 
    -\alpha_t^2\frac{(L+2c_{t+1})\Gamma\VARIS}{2B} \nonumber\\

@@

	-\frac{f_t}{B}\VARIS,\nonumber\\

@@

	-\frac{f_t}{B}\VARIS, \label{eq:20}

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\{\Psi_t\}, 
& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi}, 
&& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi},
\end{align*}
where $\Psi$, $d_t$ and $f_t$ are defined in Lemma~\ref{lemma:aux1}.
% The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ and using telescopic sum we obtain
\begin{align*}
\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
 \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 & = \frac{R^{s+1}_m - R^{s+1}_0}{\psi}  + \frac{m\zeta}{N} + \frac{m\xi}{B}
\end{align*}
By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that
\todopir{Why $c_m =0$? In the previous theorem  was $c_{m-1} = 0$}
\begin{align*}
        R^{s+1}_m - R^{s+1}_0 
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})- c_m \norm[]{\vtheta^{s+1}_{m} - \wt{\vtheta}^s}^2 } - \EVV[0|s]{J(\vtheta^{s+1}_{0}) - c_0 \norm[]{\vtheta^{s+1}_0 - \wt{\vtheta}^s}^2}\\
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})} - \EVV[0|s]{J(\wt{\vtheta}^s)}
        = \EVV[m|s]{J(\wt{\vtheta}^{s+1}) - J(\wt{\vtheta}^s)}
\end{align*}
Next, summing over epochs:
\begin{align}
\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
%
&\leq
\frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
 \label{eq:23}\\
%
&\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
\end{align}
where the expectation in~\eqref{eq:23} is \wrt all the trajectories sampled in a run of Algorithm~\ref{alg:svrpg} and~\eqref{eq:24} is from the definition of $\vtheta^*$ (\ie the policy performance maximizer).
Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
\begin{align*}
\EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}.
\end{align*}

%% \begin{align}
%% \sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}^2}}&\leq
%%  \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
%%  %
%%  &= \frac{\sum_{t=0}^{m-1}\left(\Ets[t+1]{J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2} - \Ets[t]{J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2}\right)}{\psi} \nonumber\\
%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
%%  %
%%  &= \frac{\Ets[m]{\sum_{t=0}^{m-1}\left(J(\vtheta_{t+1}^{s+1}) - c_{t+1}\norm[]{\vtheta_{t+1}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{t}^{s+1}) - c_{t}\norm[]{\vtheta_{t}^{s+1}-\tilde{\vtheta}^s}^2\right)}}{\psi} \nonumber\\
%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
%%  %
%%  &= \frac{\Ets[m]{J(\vtheta_{m}^{s+1}) - c_{m}\norm[]{\vtheta_{m}^{s+1}-\tilde{\vtheta}^s}^2 - J(\vtheta_{0}^{s+1}) - c_{0}\norm[]{\vtheta_{0}^{s+1}-\tilde{\vtheta}^s}^2}}{\psi} \nonumber\\
%%  &\qquad+ \frac{m\zeta}{N} + \frac{m\xi}{B} \label{eq:21}\\
%%  %
%%  &= \frac{\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B}, \label{eq:22}
%%  \end{align}
%% where (\ref{eq:21}) is obtained by telescoping the sum and (\ref{eq:22}) from the fact that $c_m=0$ and $\vtheta_0^{s+1}=\tilde{\vtheta}^s$.
%% Next, summing over epochs:
%% \begin{align}
%% \sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
%% \frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
%% %
%% &\leq
%% \frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
%%  \label{eq:23}\\
%% %
%% &\leq
%% \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
%% \end{align}
%% where the expectation in (\ref{eq:23}) is \wrt all the trajectories sampled in a run of Algorithm \ref{alg:svrpg} and (\ref{eq:24}) is from the definition of $\vtheta^*$.
%%  
%% Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
%% \begin{align}
%% \EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
%% &=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} \nonumber\\
%% &\leq
%% \frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B} \nonumber.
%% \end{align}
\end{proof}
